---
title: "Project Template-3"
output:
  word_document: default
  html_document: default
---


# Section 3


### Introduction. 

**Customer Segmentation** is one the most important modeling of unsupervised learning. Using clustering techniques, business can identify the several segments of customers allowing them to target the potential user base. In this machine learning project, this project will make use of K-means clustering which is the essential algorithm for clustering unlabeled dataset. 

**Source of the dataset :**  https://www.kaggle.com/vjchoudhary7/customer-segmentation-tutorial-in-python 

**Purpose : ** This data set is created only for the learning purpose of the customer segmentation concepts , also known as market basket analysis. 

**Content** 

This is a sample data to  mimics supermarket mall data which can be acquired through membership cards , It  has some basic data about regular mall customers  like : 

* Customer ID 
* Age 
* Gender 
* Annual Income and 
* Spending score 
    
    
### The problem statement 

Project targets to present customer segmentation analysis in form of visual representation to marketing team of a shopping mall to help them to plan effective marketing stretegies for prosprective customres. Based on available variables in dataset , project is trying to provide analysis on following data points :

* Price segmentation is basic and  very common factor to determine prospective customers. 
* How does demographic information like Gender/ Age can be used to identify prospective customer ? 
* How does annual Income can be used to Identify customer buying pattern ? 
* Finally spending score to determine prospective customer. 


### Approach  (How you addressed this problem statement )

Project begins with loading dataset from .csv file followed by basic statistical analysis and data exploration. Finally, going  through the input data to gain necessary insights about it. 

Following are key analysis and their presentations  :

1. Customer Gender Visualization
2. Visualization of Age Distribution
3. Analysis of the Annual Income of the Customers
4. Analyzing Spending Score of the Customers
5. Specifically,  use of a clustering algorithm called **K-means clustering** by  analyzing and visualizing the data , cluster assignment and finally determining optimal cluster using following three popular methods : 

i.   Elbow method 
ii.  Silhouette method 
iii. Gap statistic 



### Analysis 

#### Important package libraries :

```{r}
library(plotrix)

library(purrr)
library(cluster) 
library(gridExtra)
library(grid)

library(NbClust)
library(factoextra)

```

### Shopping mall dataset 

```{r }

setwd("~/MadR/Workspaces/dsc520")
customer_data=read.csv("~/MadR/Workspaces/dsc520/data/Mall_Customers.csv")

head(customer_data)
summary(customer_data)
str(customer_data)


```

### Data Exploration

```{r}

'Age standard deviation'
sd(customer_data$Age)
'Annual Income standard deviation'
sd(customer_data$Annual.Income..k..)
'Spending standard deviation'
sd(customer_data$Spending.Score..1.100.)

```


### Implications 

Analysis presented in the project helps to identify prospective customers based on their age group, Income range , gender and spending scores. This analysis was done on a smaller set of  available customer data of a shopping mall. Same model can be replicated for any other marketing team (from other businesses) to build their stretegy to target prospective cudtomers. 

However, this study was done on a limited sized data with only 5 variables.This is recommnded to collect more customer data to build a  bigger dataset with  few more  demographic information along with   shopping history of a customer for better insight.

## Visualization and Statistics

### Customer Gender Visualization 

```{r}

a=table(customer_data$Gender)
barplot(a,
        main = "Gender Comparision",
        ylab = 'Count',
        xlab = 'Gender',
        col=rainbow(2),
        legend=rownames(a) )

pct=round(a/sum(a)*100)
lbs=paste(c("Female","Male")," " ,pct,"%",sep=" ")
pie3D(a,
      labels  =lbs,
      main="Ratio of Female and Male")

```

### Customer Age  Visualization 
maximum customer ages are between 30 to 35 with minimum /maximum ages are  18 & 70 respectively

```{r}

summary(customer_data$Age)

hist(customer_data$Age,
     col="blue",
     main="Age class",
     xlab="Age Class",
     ylab  ="Frequency",
     labels=TRUE)

boxplot(customer_data$Age,
        col="#ff0066",
        main="Age")
```


###  Analyzing Annual Income
Income range lies between 15 to 137K range with average income being 60.56

```{r}

summary(customer_data$Annual.Income..k..)

hist(customer_data$Annual.Income..k,
     col="#660033",
     main="Histogram of Annual Income",
     xlab="Annual Income Class",
     ylab="Frequency",
     labels=TRUE)

plot(density(customer_data$Annual.Income..k..),
     col="yellow",
     main="Density Plot for Annual Income",
     xlab="Annual Income Class",
     ylab="Density")
polygon(density (customer_data$Annual.Income..k..),col="#ccff66")
```


### Analyzing Spending Score
Spending score range is between 1 to 99 with average of 50.20. Histogram indicates customer between 40 and 50 have the highest spending score 

```{r}


summary(customer_data$Spending.Score..1.100.)
boxplot(customer_data$Spending.Score..1.100.,
        horizontal = TRUE,
        col="#990000",
        main="Boxplot of Spending Score")

hist(customer_data$Spending.Score..1.100.,
     main = "Histogram for spending Score",
     xlab = "Spending Score Class",
     ylab = "Frequency",
     labels = TRUE)

```




### K-means Algorithm


```{r}



set.seed(123)
# function to calculate total intra-cluster sum of square 
iss <- function(k) {
  kmeans(customer_data[,3:5],k,iter.max=100,nstart=100,algorithm="Lloyd" )$tot.withinss
}

k.values <- 1:10


iss_values <- map_dbl(k.values, iss)

plot(k.values, iss_values,
    type="b", pch = 19, frame = FALSE, 
    xlab="Number of clusters K",
    ylab="Total intra-clusters sum of squares")
```



```{r}




k2<-kmeans(customer_data[,3:5],2,iter.max=100,nstart=50,algorithm="Lloyd")
s2<-plot(silhouette(k2$cluster,dist(customer_data[,3:5],"euclidean")))
```


```{r}

k3<-kmeans(customer_data[,3:5],3,iter.max=100,nstart=50,algorithm="Lloyd")
s3<-plot(silhouette(k3$cluster,dist(customer_data[,3:5],"euclidean")))


```


```{r}
k4<-kmeans(customer_data[,3:5],4,iter.max=100,nstart=50,algorithm="Lloyd")
s4<-plot(silhouette(k4$cluster,dist(customer_data[,3:5],"euclidean")))
```



```{r}
k5<-kmeans(customer_data[,3:5],5,iter.max=100,nstart=50,algorithm="Lloyd")
s5<-plot(silhouette(k5$cluster,dist(customer_data[,3:5],"euclidean")))


```

```{r}
k6<-kmeans(customer_data[,3:5],6,iter.max=100,nstart=50,algorithm="Lloyd")
s6<-plot(silhouette(k6$cluster,dist(customer_data[,3:5],"euclidean")))
```

```{r}
k7<-kmeans(customer_data[,3:5],7,iter.max=100,nstart=50,algorithm="Lloyd")
s7<-plot(silhouette(k7$cluster,dist(customer_data[,3:5],"euclidean")))
```

```{r}
k8<-kmeans(customer_data[,3:5],8,iter.max=100,nstart=50,algorithm="Lloyd")
s8<-plot(silhouette(k8$cluster,dist(customer_data[,3:5],"euclidean")))
```

```{r}
k9<-kmeans(customer_data[,3:5],9,iter.max=100,nstart=50,algorithm="Lloyd")
s9<-plot(silhouette(k9$cluster,dist(customer_data[,3:5],"euclidean")))
```

```{r}
k10<-kmeans(customer_data[,3:5],10,iter.max=100,nstart=50,algorithm="Lloyd")
s10<-plot(silhouette(k10$cluster,dist(customer_data[,3:5],"euclidean")))
```

```{r}


fviz_nbclust(customer_data[,3:5], kmeans, method = "silhouette")
```
```{r}
set.seed(125)
stat_gap <- clusGap(customer_data[,3:5], FUN = kmeans, nstart = 25,
            K.max = 10, B = 50)
fviz_gap_stat(stat_gap)

```
```{r}
k6<-kmeans(customer_data[,3:5],6,iter.max=100,nstart=50,algorithm="Lloyd")
k6
```
```{r}
pcclust=prcomp(customer_data[,3:5],scale=FALSE) #principal component analysis
summary(pcclust)

pcclust$rotation[,1:2]
```

```{r}
set.seed(1)
ggplot(customer_data, aes(x =Annual.Income..k.., y = Spending.Score..1.100.)) + 
  geom_point(stat = "identity", aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
              breaks=c("1", "2", "3", "4", "5","6"),
              labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")
```
```{r}
ggplot(customer_data, aes(x =Spending.Score..1.100., y =Age)) + 
  geom_point(stat = "identity", aes(color = as.factor(k6$cluster))) +
  scale_color_discrete(name=" ",
                      breaks=c("1", "2", "3", "4", "5","6"),
                      labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5","Cluster 6")) +
  ggtitle("Segments of Mall Customers", subtitle = "Using K-means Clustering")
```

```{r}
kCols=function(vec){cols=rainbow (length (unique (vec)))
return (cols[as.numeric(as.factor(vec))])}

digCluster<-k6$cluster; dignm<-as.character(digCluster); # K-means clusters

plot(pcclust$x[,1:2], col =kCols(digCluster),pch =19,xlab ="K-means",ylab="classes")
legend("bottomleft",unique(dignm),fill=unique(kCols(digCluster)))
```



### Limitations 

In the process of project execution , I came across following limitations :


1. Customer tend to behave differently, and think differently, at different times or occasions. For example, dietary habits and preferences
vary by occasion: Friday night diner is different from lunch during the week. Existing dataset doesn't have any variable with any behavarial attribute of the customer.

2. Shopping score can be derrived from shopping history in conjuction with shopping event but existing dataset feeds shopping score directly without any dependent variable being captured. 

3. Data size to get more accurate reading and predictions.



### Concluding Remarks

Project is  developed using unsupervised ML technique **(KMeans Clustering Algorithm)** in the simplest form. Project mimics supermarket shopping mall data which can be acquired through  membership cards , Dataset provides basic data about  customers like Customer ID, age, gender, annual income and spending score. 

K-Menas clustering helps to segment  group of customer on the basis of their age, gender, income and shopping score along with various visual presentations for marketing team to identify prospective customers to plan their strategy.

Project scope is limited to shopping mall but this approach can be reused for other domains by factoring variables of respective datasets to conduct unsupervised ML for respective marketing team.











