---
title: "Exerise14_AyachitMadhukar"
output:
  pdf_document: default
  html_document.pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Exercise 14

```{r echo=TRUE}

setwd("~/MadR/Workspaces/dsc520")


```

### Loading data from file 

```{r}
data=read.csv("data/binary-classifier-data.csv")

```

### Observations at first glance  
```{r}
#Normalization
summary(data)
str(data)
head(data)

```

### Randomization 
```{r}

set.seed(9850)
gp<-runif(nrow(data))
data<-data[order(gp),]
head(data)  

```

### Normalization 

```{r}
normalize<-function(x){
  return (
            (x - min(x))/max(x)-min(x)
         )
}
data.n<-as.data.frame(lapply(data[,c(2:3)], normalize))

str(data.n)
summary(data.n)
```
### Preparing training/testing models 

```{r}

r<-round(0.8*nrow(data.n))
l<-nrow(data.n)

data.train<-data.n[1:r,]
data.test<-data.n[r:l,]

data.train.target<-data[1:r,1]
data.test.target<-data[r:l,1]


```
## Logistic regression
```{r}

library(caTools)

split<- sample.split(data,SplitRatio=.8)

train<-subset(data,split=="TRUE")
test<-subset(data,split=="FALSE")

GLM.1 <- glm(train$label ~ ., family=binomial(), data=train)
summary(GLM.1)

res<-predict(GLM.1,test,type="response")
res<-predict(GLM.1,train,type="response")

conffmatrix<- table(actual_value=train$label,Predicted_value=res>0.5)

conffmatrix

acc<-(conffmatrix[1,1] + conffmatrix[2,2])/(conffmatrix[1,1] + conffmatrix[2,2] + conffmatrix[2,1]+ conffmatrix[1,2])*100



```

## KNN Classifire
### Determining k value 
```{r}
 k_value<-round(sqrt(nrow(data)))
```

### Predicting knn and accuracy

```{r}
library("class")

knn.39<- knn(train = data.train, test=data.test,cl=data.train.target, k=k_value)
ACC.39<-100 * sum( data.test.target == knn.39) / NROW(data.test.target)
                                                     
table(data.test.target,knn.39)

ACC.39
knn.39<- knn(train = data.train, test=data.test,cl=data.train.target, k=k_value)


```
## Conclusion 

### a. What is the accuracy of the logistic regression classifier? 
>*logistic regression shows close to `r ACC.39` accuracy*  

### b. How does the accuracy of the logistic regression classifier compare to the nearest neighbors algorithm?
>*logistic regression is way off from KNN whch is nearly `r acc`% vs KNN being `r ACC.39`% accurate* 

### c. Why is the accuracy of the logistic regression classifier different from that of the nearest neighbors?
>**** 


